---
title: "LP Problem"
author: "Osama Bin Zahir"
date: "2023-09-25"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Writing Linear Programming Problem

# Objective
# Max = 420(Al + Bl + Cl) + 360(Am + Bm + Cm) + 300(As + Bs + Cs)

# Production Constraint
# Al + Am + As <= 750
# Bl + Bm + Bs <= 900
# Cl + Cm + Cs <= 450

# Storage Constraint
# 20Al + 15Am + 12As <= 13000
# 20Bl + 15Bm + 12Bs <= 12000
# 20Cl + 15Cm + 12Cs <= 5000

# Sales Constraint
# Al + Bl + Cl <= 900
# Am + Bm + Cm <= 1200
# As + Bs + Cs <= 750

# Same Ratio Constraint
# Al + Am + As / 750 = Bl + Bm + Bs / 900 = Cl + Cm + Cs / 450

# Loading required libraries

```{r}
library(lpSolve)
```

# Coefficients and constraint matrix 

```{r}
coeff = c(420,360,300,420,360,300,420,360,300)

mat = matrix(c(1, 0, 0, 1, 0, 0, 1, 0, 0,
                 0, 1, 0, 0, 1, 0, 0, 1, 0,
                 0, 0, 1, 0, 0, 1, 0, 0, 1,
                 1, 1, 1, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 1, 1, 1, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 1, 1, 1,
                 20, 15, 12, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 20, 15, 12, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 20, 15, 12,
                 900, 900, 900, -750, -750, -750, 0, 0, 0,
                 0, 0, 0, 450, 450, 450, -900, -900, -900,
                 450, 450, 450, 0, 0, 0, -750, -750, -750),
               nrow=12,
               byrow=TRUE)
```

# Directions of constraints and RHS constraints

```{r}
dir = c("<=","<=","<=","<=","<=","<=","<=","<=","<=","=","=","=")

RHS = c(900,1200,750,750,900,450,13000,12000,5000,0,0,0)

```

# Solving the problem

```{r}
X = lp("max", coeff, mat, dir, RHS)
```

# Displaying output values

```{r}
cat("Optimal Profit (P):\n")
print(X$objval)

cat("Optimal Solution (Al, Am, As, Bl, Bm, Bs, Cl, Cm, Cs):\n")
X$solution[1]
X$solution[2]
X$solution[3]
X$solution[4]
X$solution[5]
X$solution[6]
X$solution[7]
X$solution[8]
X$solution[9]

```
# Quantity of products to be manufactured at Plant Al, Am, As:

# Plant 1:
#    Al = 516.6667
#    Am = 177.7778
#    As = 0

# Quantity of products to be manufactured at Plant Bl, Bm, Bs:

# Plant 2:

#    Bl = 0
#    Bm = 666.6667
#    Bs = 166.6667

# Quantity of products to be manufactured at Plant Cl, Cm, Cs:

# Plant 3:

#    Cl = 0
#    Cm = 0
#    Cs = 416.6667